\textbf{Using the MSU Thesis Class}

### 2022-12-14 docs/msu-thesis.tex
v4 updated documentation and template

\def\msuversion{3.7}
\def\msudate{2022/09/05}
\title{\textbf{Using the MSU Thesis Class}}
\author{\textbf{Alan Munn}\\Department of Linguistics, Languages, and Cultures\\Michigan State University\\\texttt{\href{mailto:amunn@msu.edu}{amunn@msu.edu}}}
\date{Version \msuversion\\\msudate}
\usepackage{fontspec}
\usepackage[lmargin=.75in,rmargin=.75in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{titling}
\usepackage{array, booktabs, multicol, fancyhdr, xspace,tabularx}
\usepackage{enumitem}
\usepackage{fancyvrb,listings,url}
\usepackage[sf,small]{titlesec}
\usepackage[section]{placeins}
\usepackage[colorlinks=true]{hyperref}
\setmonofont{Inconsolatazi4}

@@

\abstract{\noindent\begin{quote} This is a class file for MSU theses and dissertations.  It is based on the \pkg{memoir}\ class, and therefore supports all of the functionality of that class.  It should generate a document which meets all the basic formatting requirements laid out by the \href{https://grad.msu.edu/etd/formatting-guide}{Formatting Guide For Submission of Master’s Theses and Doctoral Dissertations}\ (March 2020) produced by the Graduate School, including requirements for margins, titlepage, page numbering, section titles, sub- and superscript sizes etc.\end{quote}}

@@

\section{User commands}
In addition to all the user functionality defined by \pkg{memoir}, the \pkg{msu-thesis}\ class defines seven new titling commands, and two “appeasement” commands, i.e., commands to tweak the formatting to appease the the thesis office’s inconsistency.
\begin{table}[htpb]
\centering
\begin{tabularx}{.8\textwidth}{>{\ttfamily}lX}
\toprule
{\bs maketitlepage} & Makes the main title page for the thesis.\\
{\bs fieldofstudy\{\}} &  Takes one argument corresponding to your field of study.\\
{\bs dedication\{\}} &  Takes one argument (should be short), your dedication.\\
{\{publicabstract\}} &  Environment for the public abstract (unnumbered, uncounted).\\
{\bs makecopyrightpage} &  Creates the copyright page (use \texttt{\bs makecopyrightpage*} to align the text on the left.)\\
{\bs makededicationpage} &  Creates the dedication page.\\
{\bs makebibliographypage} &  Creates the bibliography cover page.\\
{\{abbreviations\}} & Environment for the Key to Abbreviations \\
\bottomrule
\end{tabularx}
\caption{Class titling commands}
\end{table}

@@

Most book-length documents are divided into three main kinds of parts: the front matter, the main matter, and the back matter.  In \pkg{memoir}\ (and therefore \pkg{msu-thesis}) these sections are preceded by the commands |\frontmatter|, |\mainmatter|, and |\backmatter|.  These commands tell the class when to change the page numbering, for example, front matter pages  use lower case roman numerals, but main matter pages use arabic numerals.

@@

\begin{table}[ht]
\begin{scheme}
\item |\begin{document}|
\item |\frontmatter|
\begin{scheme}
	\item Titlepage
	\item Public Abstract
	\item Abstract
	\item Copyright Page
	\item Dedication
	\item Acknowledgements
	\item Preface
	\item Table of Contents
	\item List of Tables
	\item List of Figures
	\item List of Abbreviations/Symbols
\end{scheme}
\item |\mainmatter|
\begin{scheme}
	\item Chapter 1
	\item Chapter 2
	\item \ldots
	\item Chapter $n$
	\item{Appendices}
\end{scheme}
\item |\backmatter|
\begin{scheme}
	\item{Bibliography}
\end{scheme}
\item |\end{document}|
\end{scheme}
\caption{Basic structure of a thesis or dissertation}\label{structure}
\end{table}
\subsection{What does the class file do?}
The class itself sets up the margins, page numbering, and formatting of all of the required pieces.  It doesn't put everything in the right order, so you are responsible for that.  The best way to do this is to use the included sample \LaTeX\ file as a basic template for your thesis.
\subsection{Line Spacing}
The \pkg{memoir}\ class provides commands for single and double spacing (|\SingleSpacing| and |\DoubleSpacing|) most of your thesis will be double spaced, but certain parts (such as the table of contents, lists of figures/tales, bibliography) may be single spaced. This duplicates the functionality of the \pkg{setspace}\ package, which therefore should not be used.  For bibliography spacing issues, see also Section \ref{bibspacing}.

@@

\subsection{Key to Abbreviations}
Another common kind of list is a List of Abbreviations. The thesis office seems to want to call this “Key to Abbreviations”.  The thesis office seems to think that these should be formatted identically to other lists, even though they are not really contents lists. Because of this I have added a built-in environment  |{abbreviations}| and a helper macro to add elements to the list. The basic idea is the following: for each abbreviation/symbol you add to the list use the command |\abbrev{<abbreviation>}{<explanation>}|. This command takes two arguments, the first being the abbreviation itself, and the second the explanation of the abbreviation. For example, for a list of abbreviations containing the abbreviations MSU and MIT we would use the following code. The list is not automatically sorted, so you need to add abbreviations with |\abbrev| in exactly the order you want them to appear in the list.

\begin{quote}
\begin{lstlisting}
\begin{abbreviations}
\abbrev{MIT}{Massachusetts Institute of Technology}
\abbrev{MSU}{Michigan State University}
\end{abbreviation}
\end{lstlisting}
\end{quote}
\subsubsection{Summary of abbreviation commands}
In addition to the |\abbrev| command, there are a few other commands to change the format of the the |\listofabbreviations| as shown in Table~\ref{abbrev}. Note that the Key to Abbreviations written to the auxiliary file with extension |.lob| which can be removed along with other |.aux| files if needed.

@@

\subsubsection{Appendices per chapter}
If you need to add an appendix or appendices to a chapter, you must use the \pkg{[chapterapp]}\ class option. The code for the appendices themselves is  exactly as described above. This means that the appendix itself is introduced by the |\chapter| command, \emph{not} a |\section| command (because the this seems to be how the thesis office wants things to look).  Furthermore,  per-chapter appendices should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

\subsection{Bibliographies}
The class places no constraints on how you do your bibliography. It doesn't care whether you are using the \pkg{chapterbib}\ package with \pkg{natbib}\ or the per-chapter bibliography functionality provided by \pkg{biblatex}. If you have per-chapter bibliographies, however, you \emph{must}\ use the |[chapterbib]| class option.  This option lets the class know that you are using per-chapter bibliographies.   Before any bibliography (whether global or per-chapter) you should use the |\makebibliographypage| command. Note that per-chapter bibliographies should \emph{not} be preceded by the |\backmatter| command, which should only precede the global bibliography or appendices.

>>>>

\def\msuversion{4.0}
\def\msudate{2022/12/13}
\title{\textbf{Using the MSU Thesis Class}}
\author{\textbf{Alan Munn}\\Department of Linguistics, Languages, and Cultures\\Michigan State University\\\texttt{\href{mailto:amunn@msu.edu}{amunn@msu.edu}}}
\date{Version \msuversion\\\msudate}
\usepackage{fontspec}
\usepackage[lmargin=.75in,rmargin=.75in,tmargin=1in,bmargin=1in]{geometry}
\usepackage{titling}
\usepackage{array, booktabs, multicol, fancyhdr, xspace,tabularx}
\usepackage{enumitem}
\usepackage{fancyvrb,listings,url}
\usepackage[sf,small]{titlesec}
\usepackage[section]{placeins}
\usepackage[colorlinks=true]{hyperref}
\setmonofont{Inconsolatazi4}

@@

\abstract{\noindent\begin{quote} This is a class file for MSU theses and dissertations.  It is based on the \pkg{memoir}\ class, and therefore supports all of the functionality of that class.  It should generate a document which meets all the basic formatting requirements laid out by the \href{https://grad.msu.edu/etd/formatting-guide}{Formatting Guide For Submission of Master’s Theses and Doctoral Dissertations}\ (September, 2022) produced by the Graduate School, including requirements for margins, titlepage, page numbering, section titles, sub- and superscript sizes etc.\end{quote}}

@@

\section{User commands}
In addition to all the user functionality defined by \pkg{memoir}, the \pkg{msu-thesis}\ class defines seven new titling commands, and two “appeasement” commands, i.e., commands to tweak the formatting to appease the the thesis office’s inconsistency.
\begin{table}[htpb]
\centering
\begin{tabularx}{.8\textwidth}{>{\ttfamily}lX}
\toprule
{\bs maketitlepage} & Makes the main title page for the thesis.\\
{\bs fieldofstudy\{\}} &  Takes one argument corresponding to your field of study.\\
{\bs dedication\{\}} &  Takes one argument (should be short), your dedication.\\
{\{publicabstract\}} &  Environment for the public abstract (unnumbered, uncounted).\\
{\bs makecopyrightpage} &  Creates the copyright page (use \texttt{\bs makecopyrightpage*} to align the text on the left.)\\
{\bs makededicationpage} &  Creates the dedication page.\\
%{\bs makebibliographypage} &  Creates the bibliography cover page.\\
{\{abbreviations\}} & Environment for the List of Abbreviations \\
\bottomrule
\end{tabularx}
\caption{Class titling commands}
\end{table}

@@

Most book-length documents are divided into three main kinds of parts: the front matter, the main matter, and the back matter.  In \pkg{memoir}\ (and therefore \pkg{msu-thesis}) these sections are preceded by the commands |\frontmatter|, and |\mainmatter|. The way the thesis office now structures the document means that the  |\backmatter| should not be used.  These commands tell the class when to change the page numbering, for example, front matter pages  use lower case roman numerals, but main matter pages use arabic numerals.

@@

\begin{table}[ht]
\begin{scheme}
\item |\begin{document}|
\item |\frontmatter|
\begin{scheme}
	\item Titlepage
	\item Public Abstract
	\item Abstract
	\item Copyright Page
	\item Dedication
	\item Acknowledgements
	\item Preface
	\item Table of Contents
	\item List of Abbreviations/Symbols
\end{scheme}
\item |\mainmatter|
\begin{scheme}
	\item Chapter 1
	\item Chapter 2
	\item \ldots
	\item Chapter $n$
\end{scheme}
\begin{scheme}
	\item{Bibliography}
\end{scheme}
\begin{scheme}
	\item{Appendices}
\end{scheme}
\item |\end{document}|
\end{scheme}
\caption{Basic structure of a thesis or dissertation}\label{structure}
\end{table}
\subsection{What does the class file do?}
The class itself sets up the margins, page numbering, and formatting of all of the required pieces.  It doesn't put everything in the right order, so you are responsible for that.  The best way to do this is to use the included sample \LaTeX\ file as a basic template for your thesis.
\subsection{Line Spacing}
The \pkg{memoir}\ class provides commands for single and double spacing (|\SingleSpacing| and |\DoubleSpacing|) most of your thesis will be double spaced, but certain parts (such as the table of contents, lists of figures/tales, bibliography) may be single spaced. This duplicates the functionality of the \pkg{setspace}\ package, which therefore should not be used.  For bibliography spacing issues, see also Section \ref{bibspacing}.

@@

\subsection{List of Abbreviations}
Another common kind of list is a List of Abbreviations.  The thesis office seems to think that these should be formatted identically to other lists, even though they are not really contents lists. Because of this I have added a built-in environment  |{abbreviations}| and a helper macro to add elements to the list. The basic idea is the following: for each abbreviation/symbol you add to the list use the command |\abbrev{<abbreviation>}{<explanation>}|. This command takes two arguments, the first being the abbreviation itself, and the second the explanation of the abbreviation. For example, for a list of abbreviations containing the abbreviations MSU and MIT we would use the following code. The list is not automatically sorted, so you need to add abbreviations with |\abbrev| in exactly the order you want them to appear in the list.

\begin{quote}
\begin{lstlisting}
\begin{abbreviations}
\abbrev{MIT}{Massachusetts Institute of Technology}
\abbrev{MSU}{Michigan State University}
\end{abbreviation}
\end{lstlisting}
\end{quote}
\subsubsection{Summary of abbreviation commands}
In addition to the |\abbrev| command, there are a few other commands to change the format of the the |\listofabbreviations| as shown in Table~\ref{abbrev}. Note that the List of Abbreviations written to the auxiliary file with extension |.lob| which can be removed along with other |.aux| files if needed.

@@

\subsubsection{Appendices per chapter}
If you need to add an appendix or appendices to a chapter, you must use the \pkg{[chapterapp]}\ class option. The code for the appendices themselves is  exactly as described above. This means that the appendix itself is introduced by the |\chapter| command, \emph{not} a |\section| command (because the this seems to be how the thesis office wants things to look).  Furthermore,  per-chapter appendices should \emph{not} be preceded by the |\backmatter| command, which should not be used at all.

\subsection{Bibliographies}
The class places no constraints on how you do your bibliography. It doesn't care whether you are using the \pkg{chapterbib}\ package with \pkg{natbib}\ or the per-chapter bibliography functionality provided by \pkg{biblatex}. If you have per-chapter bibliographies, however, you \emph{must}\ use the |[chapterbib]| class option.  This option lets the class know that you are using per-chapter bibliographies.    Note that per-chapter bibliographies should \emph{not} be preceded by the |\backmatter| command, which should not be used at all.
